---
title: "spmR canonical examples"
vignette: >
  %\VignetteIndexEntry{spmR canonical examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE)
```

```{r setup}
library(spmR)
library(readr)
```

This vignette is intentionally limited to two canonical example directories that are
kept in sync with the package:

- `examples/atka` for `runSPM()`, `dat2list()`, and `plotSPMx()`
- `examples/BSRE_AI` for `plotSPM()` with `spm_summary.csv`

## 1. Atka workflow (`examples/atka`)

Use `runSPM()` to read existing ADMB output (`spm_detail.csv`) and inspect inputs
using `dat2list()`.

```{r atka-read}
pkg_root <- if (file.exists("DESCRIPTION")) "." else ".."
atka_dir <- file.path(pkg_root, "examples", "atka")

atka_detail <- runSPM(atka_dir, run = FALSE, engine = "admb")
str(atka_detail)

atka_inputs <- dat2list(file.path(atka_dir, "spm.dat"))
names(atka_inputs)
```

Plot detailed simulation trajectories with `plotSPMx()`.

```{r atka-plot}
plotSPMx(atka_detail, alt = 2, thisyr = min(atka_detail$Year), mytitle = "Atka detail plot")
```

The experimental RTMB path can also be run from this directory when `RTMB` is
installed.

```{r atka-rtmb, eval = FALSE}
runSPM(atka_dir, run = TRUE, engine = "rtmb")
```

## 2. Summary workflow (`examples/BSRE_AI`)

`plotSPM()` expects summary-format data (`spm_summary.csv`).

```{r bsre-summary}
bsre_dir <- file.path(pkg_root, "examples", "BSRE_AI")
bsre_summary <- read_csv(file.path(bsre_dir, "spm_summary.csv"))
plotSPM(bsre_summary, alt = c(1, 3, 5, 7), mytitle = "BSRE_AI summary plot")
```
