---
title: "spmR-examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{spmR-examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setupex}
library(spmR)
library(tidyverse)
library(patchwork)
here::here()

```
# Working examples of the spmR package

This vignette provides examples of how to use the spmR package to run the Standard Projection Model (SPM) for Alaska groundfish. The SPM is a stock assessment model used to project the future status of fish stocks. The spmR package provides an R interface to the SPM model, allowing users to run the model and analyze the results in R. There are three main functions in the spmR package: `runSPM`, `plotSPM`, and `dat2list` (and `list2dat`). The `runSPM` function runs the SPM model and returns the results as a data frame. The `plotSPM` function creates plots of the SPM results. 
The `list2dat` and `dat2list` functions convert between local R lists and datafiles used by SPM externally.
A draft function called `print_Tier3_tables` is set up to provide some capabilities to create
tables for markdown documents.
This vignette demonstrates how to use these functions to run the SPM model and analyze the results. 

# Example sets

The following case studies have been applied.

## Atka mackerel
```{r atka,eval=TRUE}
df <- runSPM(here::here("examples","atka"),run=FALSE)
#setwd("spmR")
plotSPM(df)
```


## BSAI northern rock sole
```{r bsaiNRS,eval=FALSE}
df <- runSPM(here::here("examples","BSAI NRS"),run=FALSE)
plotSPM(df)
```
## EBS pollock
```{r ebswp,eval=FALSE}
df <- runSPM(here::here("examples","ebspollock"),run=FALSE)
plotSPM(df)

```
## BSAI yellowfin sole
```{r fathead,eval=FALSE}
df <- runSPM(here::here("examples","yfs"),run=FALSE)
 plotSPM(df)

```
## GOA Rougheye blackspotted
```{r rougheye,eval=FALSE}

```
