---
title: "RTMB vs ADMB comparison (experimental)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RTMB vs ADMB comparison (experimental)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This vignette compares the legacy ADMB workflow (`spm.tpl`) with the new
RTMB prototype. Its archived outputs illustrate the interface; they are
placeholders for population dynamics and cannot establish scientific equivalence.
Version 0.4.1 stops new RTMB projections until the required dynamics and input
contract are implemented. New ADMB runs require input format 2 metadata and a
male spawning-weight vector `wt_M` after `wt_F` in split-sex species files.
The [split-sex input migration guide](split_sex_inputs.html) provides a runnable
example with explicit historical recruitment basis, separate female and male
population weights, and starting-year checks. Add those fields to
`spm_metadata.json`. Version 0.4.1 no longer reads `tacpar.dat`.
Archived RTMB output remains a record of the earlier prototype; adding metadata
leaves its results and scientific limitations unchanged.

## Read archived prototype output

```{r}
# Replace with your run directory that contains spm.dat
run_dir <- c("../examples/atka", "examples/atka")
run_dir <- run_dir[file.exists(run_dir)][1]
if (is.na(run_dir)) stop("Could not locate examples/atka")
run_dir <- normalizePath(run_dir, winslash = "/", mustWork = TRUE)

set.seed(123)

# ADMB run (requires a format-2 binary and validated spm_metadata.json)
# admb_res <- runSPM(run_dir, run = TRUE, engine = "admb")

# RTMB run
rtmb_file <- file.path(run_dir, "spm_detail_rtmb.csv")
if (file.exists(rtmb_file)) {
  rtmb_res <- utils::read.csv(rtmb_file)
} else {
  stop("Need the archived spm_detail_rtmb.csv example file.")
}
```

## Compare outputs

```{r}
# If ADMB results are available, compare the distributions
# admb_res <- readr::read_csv(file.path(run_dir, "spm_detail.csv"))

if (exists("admb_res")) {
  summary(admb_res$SSB)
}
summary(rtmb_res$SSB)
```

```{r}
if (exists("admb_res")) {
  library(ggplot2)
  metrics <- c("SSB", "ABC", "OFL", "Catch", "F")
  plot_data <- rbind(
    transform(admb_res[, metrics], engine = "ADMB"),
    transform(rtmb_res[, metrics], engine = "RTMB")
  )

  plot_long <- tidyr::pivot_longer(
    plot_data,
    cols = all_of(metrics),
    names_to = "metric",
    values_to = "value"
  )

  ggplot(plot_long, aes(x = value, color = engine)) +
    geom_density() +
    facet_wrap(~metric, scales = "free") +
    labs(x = NULL, y = "Density", color = "Engine") +
    theme_minimal()
}
```

## Notes

- Archived RTMB output is read from `spm_detail_rtmb.csv`.
- Its differences from ADMB include missing population dynamics.
- Use the native regression suite for validated pooled-sex and split-sex comparisons.
