---
title: "Alaska groundfish projection model"
author: "Jim Ianelli"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	message = TRUE,
	warning = TRUE,
	collapse = TRUE,
	comment = "#>"
)
```

The projection model and methods have been used since 1999 and were designed to provide the needed projection scenarios for the annual Environmental Assessment (EA) for the TAC specifications.  
In 2005 some enhancements were made on the standardized projection model (SPM) methodology.  
to determine whether a stock is currently in an overfished condition or is approaching an overfished condition.  

```{r main, include = FALSE}
#' # Projection model work
#' Projection model work for demonstrating application of controls and input data

library(tidyverse)
library(ggplot2)
library(ggthemes)
library(gtable)
source("../R/readData.R")

#' ## Set initial "setup" parameters
thisyr=2019
setup<-list(
  Run_name     = noquote("Std"),
  Tier         = 3    ,
  nalts        = 7    ,
  alts         = c(1,2,3,4,5,6,7),
  tac_abc      = 1,    #' Flag to set TAC equal to ABC (1 means true, otherwise false)
  srr          = 1 ,   #' Stock-recruitment type (1=Ricker, 2=Bholt)
  rec_proj     = 1,    #' projection rec form (default: 1 = use observed mean and std, option 2 = use estimated SRR and estimated sigma R)
  srr_cond     = 0 ,   #' SR-Conditioning (0 means no, 1 means use Fmsy == F35%?, 2 means Fmsy == F35% and Bmsy=B35%  condition (affects SRR fits)
  srr_prior    = 0.0,  #' Condition that there is a prior that mean historical recruitment is similar to expected recruitment at half mean SSB and double mean SSB 0 means don't use, otherwise specify CV
  write_big    = 1,    #' Flag to write big file (of all simulations rather than a summary, 0 means don't do it, otherwise do it) Write_Big
  nyrs_proj    = 14,   #' Number of projection years
  nsims        = 100, #' Number of simulations
  beg_yr_label = thisyr  #' Begin Year
)

#' ## Set up the species specific run file
config<-list(
  nFixCatchYrs = 2,
  nSpecies     = 1,
  OYMin        = .1343248,
  OYMax        = 1943248,
  dataFiles    = noquote("data/t1.dat"),
  ABCMult      = 1,
  PoplnScalar  = 1000,
  AltFabcSPR   = 0.75,
  nTAC         = 1,
  TACIndices   = 1,
  Catch        = c( 2016,55000.,   2017,55000. )
)

#' ##  Make list of main file w/ assessment model results
#' E.g., "data/bsai_atka.dat"
datfile <- list(
  runname     = noquote("M16.2"), 
  ssl_spp     = 1,         # SSL_spp
  Dorn_buffer = 1,         # Dorn_buffer
  nfsh        = 1,         # N_fsh
  nsex        = 1,         # N_sexes
  avgF5yr     = 0.0661399, # avg_5yr_F
  F40_mult    = 1,         # F_40_multiplier
  spr_abc     = 0.4,       # SPR_abc
  spr_msy     = 0.35,      # SPR_msy
  sp_mo       = 8,         # spawn_month
  nages       = 11,        # N_ages
  Frat        = 1,         # F_ratio
	# M
	M    = c(0.3,0.3,0.3,0.3,0.3,0.3,0.3,0.3,0.3,0.3,0.3), 
	#  Maturity
	pmat = c(0.005,0.037,0.224,0.688,0.944,0.992,0.999,1,1,1,1),
	#  Wt_at_age_spawners  
	wtage_sp  = c(44.8,161.377,398.272,557.695,652.113,719.573,863.744,948.744,921.397,885.912,1069.87),
	#  Wt_at_age_fsh
	wtage_fsh = c(69.3778,253.522,408.211,614.731,668.483,718.137,803.017,798.707,788.117,842.468,960.006), 
	# select
	sel = c(0.002576427,0.040030753,0.651104228,0.768404263,0.794886081,1,0.889293108,0.604815671,0.451169778,0.403195516,0.403195516),
	# N
	N   = c(511.179,378.528,278.443,194.385,183.423,45.5404,61.1188,19.8073,39.6285,33.6501,51.7806),
	# Nyrs
	nyrs = 37,
	# recruits
	R    = c(1578.51,479.509,357.919,443.588,318.981,413.125,514.351,600.987,536.301,692.34,452.279,1618.73,702.801,372.811,597.812,1136.19,402.862,424.179,1025.36,207.05,383.695,1054.64,2224.52,1379.34,1545.81,345.556,454.884,617.194,404.876,993.497,727.754,236.795,505.948,258.653,726.627,524.198,473.54),
	# SSB 
	SSB  = c(206.391,194.569,187.097,183.296,195.289,240.774,252.143,238.163,223.199,200.776,182.378,179.671,189.193,198.242,212.123,233.484,277.891,282.004,250.709,231.816,218.403,195.275,181.628,189.976,175.912,168.624,220.206,315.124,376.621,397.171,365.476,317.159,277.777,242.719,233.415,223.636,198.117,183.537,177.91)
	)

#' ## Save lists for running model to files expected by projection model
# Setup.dat
list2dat(setup,"setup.dat")
# spp_catch.dat
list2dat(config,"data/t1_spcat.dat")
runfn<-"t1"
file.copy(paste0("data/",runfn,"_spcat.dat"),"spp_catch.dat",overwrite=TRUE)

list2dat(datfile,"data/t1.dat")

#' ## Run projection model
system("../src/main")
#' ## Read in projection model mainfiles
  .projdir="t1/"
  dir.create(.projdir)
  file.copy(list.files(getwd(), pattern="out$"), .projdir,overwrite=TRUE)      
  file.remove(list.files(getwd(), pattern="out$"))
  bf <- data.frame(read.table(paste0(.projdir,"bigfile.out"),header=TRUE,as.is=TRUE))
  bfs <- bf %>% filter(Sim<=30)
  #write.csv(bfs,"data/proj.csv")
 # head(bfs)
  bfss <- bfs %>% filter(Alt==2) %>% select(Alt,Yr,Catch,SSB,Sim) 
  pf <- data.frame(read.table(paste0(.projdir,"percentdb.out"),header=F) )
  names(pf) <- c("stock","Alt","Yr","variable","value") 
#' ## Make plot of projection model simulations
  p1 <- pf %>% filter(substr(variable,1,1)=="C",variable!="CStdn",Alt==2) %>% select(Yr,variable,value) %>% spread(variable,value) %>%
    ggplot(aes(x=Yr,y=CMean),width=1.2) + geom_ribbon(aes(ymax=CUCI,ymin=CLCI),fill="goldenrod",alpha=.5) + theme_few() + geom_line() +
    scale_x_continuous(breaks=seq(thisyr,thisyr+14,2))  +  xlab("Year") + ylab("Tier 3 ABC (kt)") + geom_point() + 
    expand_limits(y=0) +
    geom_line(aes(y=Cabc)) + geom_line(aes(y=Cofl),linetype="dashed") + geom_line(data=bfss,aes(x=Yr,y=Catch,col=as.factor(Sim)))+ guides(size=FALSE,fill=FALSE,alpha=FALSE,col=FALSE) 
  p2 <- pf %>% filter(substr(variable,1,1)=="S",variable!="SSBStdn",Alt==2) %>% select(Yr,variable,value) %>% spread(variable,value) %>%
    ggplot(aes(x=Yr,y=SSBMean),width=1.2) + geom_ribbon(aes(ymax=SSBUCI,ymin=SSBLCI),fill="coral",alpha=.5) + theme_few() + geom_line() +
    scale_x_continuous(breaks=seq(thisyr,thisyr+14,2))  +  xlab("Year") + ylab("Tier 3 Spawning biomass (kt)") + geom_point() + 
    expand_limits(y=0) +
    geom_line(aes(y=SSBFabc)) + geom_line(aes(y=SSBFofl),linetype="dashed")+ geom_line(data=bfss,aes(x=Yr,y=SSB,col=as.factor(Sim)))+ guides(size=FALSE,fill=FALSE,alpha=FALSE,col=FALSE) 
  t3 <- grid.arrange(p1, p2, nrow=2)
  ggsave(paste0(.projdir,"tier3_proj.pdf"),plot=t3,width=5.4,height=7,units="in")


#' ## Make tables
  # Stock Alt Sim Yr  SSB Rec Tot_biom SPR_Implied F Ntot Catch ABC OFL AvgAge AvgAgeTot SexRatio FABC FOFL
  bfsum <- bf %>% select(Alt,Yr,SSB,F,ABC ,Catch) %>% group_by(Alt,Yr) %>% summarise(Catch=mean(Catch),SSB=mean(SSB),F=mean(F),ABC=mean(ABC))
  t1 <- bfsum %>% select(Alt,Yr,Catch) %>% spread(Alt,Catch) 
  names(t1) <- c("Catch","Scenario 1","Scenario 2","Scenario 3","Scenario 4","Scenario 5","Scenario 6","Scenario 7")

  print_Tier3_tables(bf)
```



## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
